\chapter{Evaluation}\label{chap:evaluation}


These are some suggestions on what you may report on:
\begin{itemize}
  \item Benchmark Selection: Which data set did you choose to evaluate your approach on? Why did you make that specific selection? Would there have been alternatives?
  \item Hardware setup: What hardware was used (processor, RAM, etc.), and which operating systems and software were used?
  \item Results: Report the plain data (plots, tables, etc.)\ and their \emph{interpretation}, i.e., did the approach work well or not? Do we know on which subset it worked well (or badly) and why was it like this? Do the results raise further questions and thus directions for future research/investigations?
\end{itemize}

When reporting your results using graphs and plots, make sure to provide all information necessary to interpret the data, e.g., axis and graph labeling (cf.~Figure~\ref{fig:xkcd}).


\section{Empirical Evaluation}
To prove that our technique is beneficial, we conduct a standard empirical evaluation on PO domains and compare the runtime of PO planners vs. transformation + TO planners

We should also do an evaluation that shows how often the criterion works, i.e., in how many instances (per domain) we can say 'yes: translate!'

%\begin{tabular}{lcccccl}\toprule
%	& \multicolumn{3}{c}{$\tol=\tols$} & \multicolumn{3}{c}{$\tol=\told$}
%	\\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
%	& $mv$  & Rel.~err & Time    & $mv$  & Rel.~err & Time\\\midrule
%	trigmv    & 11034 & 1.3e-7 & 3.9 & 15846 & 2.7e-11 & 5.6 \\
%	trigexpmv & 21952 & 1.3e-7 & 6.2 & 31516 & 2.7e-11 & 8.8 \\
%	trigblock & 15883 & 5.2e-8 & 7.1 & 32023 & 1.1e-11 & 1.4e1\\
%	expleja   & 11180 & 8.0e-9 & 4.3 & 17348 & 1.5e-11 & 6.6 \\\bottomrule
%\end{tabular}


% https://ojs.aaai.org/index.php/ICAPS/article/view/15959/15770
% HyperTension
% Pand Pro Pand SaT GBFS Add (th, s, th+s)



\subsection{Hardware Setup}
8GB RAM
Intel(R) Core(TM) i7-10510U CPU @ 1.80GHz  (Total Cores 4; Total Threads 8)
Ubuntu 20.04.4 LTS
C++

\subsection{Solving Performance}
For this benchmark we will be doing
\begin{tabular}{lcccl} 
	\multicolumn{5}{c} {Rover} \\\toprule 
	Tasks &  time to solve PO &  time to solve TO &  PO/TO time &  Problem Name \\
	396 & 203.617 & 0.168 & 121032.0 & Rover/pfile12 \\ 
	460 & 187.619 & 4.077 & 4601.0 & Rover/pfile10 \\ 
	65 & 0.236 & 0.057 & 410.0 & Rover/pfile01 \\ 
	59 & 0.111 & 0.055 & 200.0 & Rover/pfile02 \\ 
	1202 & 380.662 & 5.17 & 7363.0 & Rover/pfile17 \\ 
	382 & 155.551 & 0.6 & 25923.0 & Rover/pfile08 \\ 
	561 & 148.941 & 0.191 & 77783.0 & Rover/pfile14 \\ 
	626 & 204.075 & 4.275 & 4773.0 & Rover/pfile16 \\ 
	89 & 0.134 & 0.058 & 228.0 & Rover/pfile03 \\ 
	595 & 202.559 & 0.403 & 50291.0 & Rover/pfile15 \\ 
	494 & 179.874 & 24.747 & 726.0 & Rover/pfile11 \\ 
	245 & 125.468 & 0.257 & 48861.0 & Rover/pfile06 \\ 
	152 & 134.515 & 0.152 & 88698.0 & Rover/pfile05 \\ 
	414 & 201.005 & 0.204 & 98302.0 & Rover/pfile09 \\ 
	707 & 261.94 & 47.815 & 547.0 & Rover/pfile13 \\\bottomrule
\end{tabular}
